---
title: "Prediction of preschool education enrollment in Central Asian countries"
subtitle: "Data Report"
author: "Sofiya Berdiyeva"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: true
    highlight: tango
---
```{=html}
<style>
.scroll-box {
  max-height: 400px;
  overflow-y: auto;
  border: 1px solid #ddd;
  padding: 6px;
  margin-bottom: 1em;
}
</style>
```

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
source(here::here("R", "00_load_packages.R"))
source(here::here("R", "01_data_cleaning.R"))

```

#  1. Loading the preprocessed data

Pre-processing included the following steps *(in the back end)*:

1. Opening SPSS (.sav) files for children under 5 questionnaire, for households and for household members

2. Removing all variables that are irrelevant for the analysis

3. Filtering out children outside the range of 3-4 years old

4. Adding unique indexes to each of the dataframes based on the logic:

"region_id" + "_" + "cluster_id" + "_" + "household_id" + "_" + "line number"

("+" = concatenation)

5. (Left) joining dataframes to retrieve all necessary variables from each questionnaire

6. Removal of duplicating variables after joins

```{r}
kazakhstan <- get_and_preprocess_mics_data("Kazakhstan")
kyrgyzstan <- get_and_preprocess_mics_data("Kyrgyzstan")
turkmenistan <- get_and_preprocess_mics_data("Turkmenistan")
uzbekistan <- get_and_preprocess_mics_data("Uzbekistan")

tajikistan <- preprocess_dhs_data(child_set_path = "data/raw/Tajikistan/TJKR81SV_childrens_recode/TJKR81FL.sav",
                                  house_members_path = "data/raw/Tajikistan/TJPR81SV_household_member_recode/TJPR81FL.sav")
```

## Kazakhstan

```{r, results = "asis"}
cat('<div class="scroll-box">')
view_df(kazakhstan, show.na = TRUE, show.frq = TRUE, show.prc = TRUE)
cat('</div>')
```

## Kyrgyzstan

```{r, results = "asis"}
cat('<div class="scroll-box">')
view_df(kyrgyzstan, show.na = TRUE, show.frq = TRUE, show.prc = TRUE)
cat('</div>')
```

## Turkmenistan

```{r, results = "asis"}
cat('<div class="scroll-box">')
view_df(turkmenistan, show.na = TRUE, show.frq = TRUE, show.prc = TRUE)
cat('</div>')
```

## Uzbekistan

```{r, results = "asis"}
cat('<div class="scroll-box">')
view_df(uzbekistan, show.na = TRUE, show.frq = TRUE, show.prc = TRUE)
cat('</div>')
```

## Tajikistan

```{r, results = "asis"}
cat('<div class="scroll-box">')
view_df(tajikistan, show.na = TRUE, show.frq = TRUE, show.prc = TRUE)
cat('</div>')
```

# 2. Outcome Variable(s)
# 3. Covariates
# 4. Missing Data
# 5. Cross-country Comparison
# 6. Implications for Empirical Analysis

# 7. Summary: concerns to discuss

