---
title: "Kyrgyzstan MICS 2023"
author: "Berdiyeva Sofiya"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: true
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache=T)
htmlwidgets::setWidgetIdSeed(123)

source(here::here("R", "00_load_packages.R"))
source(here::here("R", "01_data_cleaning.R"))
source(here::here("R", "02_descriptives.R"))
set.seed(42)
```

# Loading data

```{r}
# Indiv-level data
kyrgyzstan_mics <- get_and_preprocess_mics_data("Kyrgyzstan")

# Region-level data
region_kg <- read_csv("../data/raw/regional_data_Kyrgyzstan_2023.csv")
```

# Preprocessing individual level data 

## Dimensionality reduction

```{r}
kyrgyzstan_mics <- kyrgyzstan_mics %>% 
  
  # Learning materials
  recode_playthings(books_var = "EC1",
                    playthings_vars = c("EC2A", "EC2B", "EC2C")) %>% 
  
  # Inadequate supervision
  recode_supervision(supervision_vars = c("EC3A", "EC3B")) %>% 
  
  # Child discipline - psychological aggression
  recode_psy_aggression(var1 = "UCD2D", var2 = "UCD2H") %>% 
  
  # Early stimulation and responsive care
  recode_early_stimulation_v2(
    activity_prefixes = c("EC5A", "EC5B", "EC5C", "EC5D", "EC5E", "EC5F")) %>% 
  
  # ECDI
  recode_ecdi_and_track(age_col = "CAGE",
                        ec_regular = paste0("EC", 21:38),
                        ec_frequency = "EC39",
                        ec_comparison = "EC40") %>% 
  mutate(UB8=if_else(UB8==1, 1,0)) %>% 
  
  # Social transfers
  recode_economic_assistance(unit_base = "ST4U$",
                                          number_base = "ST4N$",
                                          n_programs = 5,
                                          months_threshold = 3) %>% 
  
  # Exclude children under 3
  filter(CAGE>35) %>% 
  
  # Sample one child from household
  sample_one_per_group(group_var = "house_index",
                       seed = 42) %>% 

  # Omit rows with missing values for key predictors
  filter_at(vars(ED5A, HL13, HL17), all_vars(!is.na(.)))
```

## Imputation

```{r}
kyrgyzstan_mics$UB8[is.na(kyrgyzstan_mics$UB8)] <- 0
```


# Correlations and associations

```{r, warning=F}

results <- pairwise_chi_cramers_parallel(
  kyrgyzstan_mics[, names(kyrgyzstan_mics) %in% c("UB2", "ED5A","UB8","HH48", "HH51","HC10E", "HH7", "HH6", "HL13", "HL17", "windex5", "ED4", "playthings_sum", "two_plus_playthings", "supervised_inadequately", "psy_aggression", "early_stimulation", "ecdi_sum", "dev_on_track", "received_econ_assist" )], n_cores = 4, p_threshold = 0.05, use_bonferroni = FALSE)


# Get summary
summary_stats(results)

# Create heatmap
plot <- plot_cramers_heatmap(results, text_size_x = 11, text_size_y = 11)
print(plot)

# View top associations
top_20 <- get_top_associations(results, n = 20)
print(top_20)
```

# Combining data for multilevel modeling

```{r}
# If we decide to grand-center anything
region_kg <- region_kg %>%
  # Log-transform to compress the scale
  mutate(popul_mln_total_01012023 = popul_thsnd_total_01012023/1000)
region_kg <- grand_mean_centering(region_kg, colnames(region_kg)[-c(1:2)])
region_kg <- region_kg %>%
  # Log-transform to compress the scale
  mutate(popul_density_log = log(popul_density),
         total_pre_2023_log = log(total_pre_2023),
         expen_on_pre_2023_log = log(expen_on_pre_2023))

kgz_df <- kyrgyzstan_mics %>%
  mutate(region_mics = as.character(as_factor(HH7))) %>% 
  left_join(region_kg,
            by = "region_mics",
            keep = FALSE)
```

# Modeling

## Baseline models

### Null model

```{r}
# RANDOM INTERCEPT WITH FIXED SLOPES

# Null model (unconditional model) - to calculate ICC
model_null <- glmer(
  UB8 ~ 1 + (1 | region_mics),
  data = kgz_df,
  family = binomial(link = "logit"),
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000))
)

# Calculate Intraclass Correlation Coefficient (ICC)
# ICC tells us how much variance is due to clustering
# An ICC > 0.05 or 0.10 or typically justifies using a multilevel model
vpc_null <- performance::icc(model_null)
print(paste("ICC:", round(vpc_null$ICC_adjusted, 3)))
```

### All individual predictors + universal regional predictors

```{r}
# Baseline random intercept model
model_for_pooling <- glmer(
  UB8 ~ 
    # Level 1 predictors (child/household)
    CAGE + # Age in months
    ED5A + # mother's/caregiver's education
    EC1 + #Three or more books
    two_plus_playthings +
    supervised_inadequately +
    psy_aggression +
    early_stimulation +
    dev_on_track +
    HH6 + # Rural/urban
    HH48 + # household size
    HL13 + # Natural mother lives in the household
    HL17 + # Natural father lives in the household
    windex5 +
    
    # Level 2 predictors (structure)
    popul_mln_total_01012023 +
    popul_density_log +
    total_pre_2023_log +

    # Random intercept for region
    (1 | region),
  
  data = kgz_df,
  family = binomial(link = "logit"),
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000))
)

# Model summary
summary(model_for_pooling, correlation =TRUE)

or_model <- model_for_pooling
ame_model <- avg_slopes(model_for_pooling, type = "response")

modelsummary(
  list(
    "Odds Ratios" = or_model,
    "Marginal Effects" = ame_model
  ),
  exponentiate = c(TRUE, FALSE),
  statistic = "conf.int",
  stars = TRUE,
  output = "gt"
)
```

```{r}
anova(model_null, model_for_pooling)
check_collinearity(model_for_pooling)
r2(model_for_pooling)
r2(model_null)
```

```{r}
# Robustness check
model_fe_region <- glm(
  UB8 ~ CAGE + # Age in months
    ED5A + # mother's/caregiver's education
    EC1 + #Three or more books
    two_plus_playthings +
    supervised_inadequately +
    psy_aggression +
    early_stimulation +
    dev_on_track +
    HH6 + # Rural/urban
    HH48 + # household size
    HL13 + # Natural mother lives in the household
    HL17 + # Natural father lives in the household
    windex5 +
    
    # Level 2 predictors (structure)
    popul_mln_total_01012023 +
    popul_density_log +
    total_pre_2023_log +
    factor(region),
  data = kgz_df,
  family = binomial
)

modelsummary(
  list(
    "Random Intercept" = model_for_pooling,
    "Region Fixed Effects" = model_fe_region
  ),
  exponentiate = TRUE
)
```

### All available country-specific variables, fixed slopes

```{r}
model_full <- glmer(
    UB8 ~ 
    # Level 1 predictors (child/household)
    CAGE + # Age in months
    ED5A + # mother's/caregiver's education
    EC1 + #Three or more books
    two_plus_playthings +
    supervised_inadequately +
    psy_aggression +
    early_stimulation +
    dev_on_track +
    received_econ_assist + # Social transfers!
    HH6 + # Rural/urban
    HH48 + # household size
    HL13 + # Natural mother lives in the household
    HL17 + # Natural father lives in the household
    windex5 +
    
    # Level 2 predictors (structure)
    popul_mln_total_01012023 +
    popul_density_log +
    total_pre_2023_log +
    expen_on_pre_2023_log +   
    employment_rate_female_2023_c +
    employment_rate_male_2023_c +
    
    
    # Random intercept for region
    (1 | region),
  
  data = kgz_df,
  family = binomial(link = "logit"),
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000))
)

summary(model_full, correlation = TRUE)

or_model_full <- model_full
ame_model_full <- avg_slopes(model_full, type = "response")
modelsummary(
  list("Odds Ratios" = or_model_full, "Marginal Effects" = ame_model_full),
  exponentiate = c(TRUE, FALSE),
  statistic = "conf.int",
  stars = TRUE,
  output = "gt"
)

anova(model_for_pooling, model_full)
check_collinearity(model_full)
r2(model_full)
```


## Reduced models
### Only significant fixed effects (p < .05) from model_full

```{r}
model_sign <- glmer(
    UB8 ~ 
    # Level 1 predictors (child/household)
    CAGE + # Age in months
    ED5A + # mother's/caregiver's education
    EC1 + #Three or more books
    HL13 +
    windex5 +
    
    # Level 2 predictors (structure)
    popul_mln_total_01012023 +
    total_pre_2023_log +

    # Random intercept for region
    (1 | region),  
  data = kgz_df,
  family = binomial(link = "logit"),
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000))
)

summary(model_sign, correlation = TRUE)

or_model_sign <- model_sign
ame_model_sign <- avg_slopes(model_sign, type = "response")
modelsummary(
  list("Odds Ratios" = or_model_sign, "Marginal Effects" = ame_model_sign),
  exponentiate = c(TRUE, FALSE),
  statistic = "conf.int",
  stars = TRUE,
  output = "gt"
)

anova(model_full, model_sign)
check_collinearity(model_sign)
r2(model_sign)
```

### Extra model: interactions + random slopes

```{r}
model_extra <- glmer(
  UB8 ~ 
    # Level 1 predictors (child/household)
    CAGE + # Age in months
    ED5A + # mother's/caregiver's education
    EC1 + #Three or more books
    HL13 +
    windex5 +
    
    # Level 2 predictors (structure)
    popul_mln_total_01012023 +
    total_pre_2023_log +
    
    # Interactions
    ED5A : employment_rate_female_2023_c +
    # Random slopes
    (1 + windex5 | region_mics),
  
  data = kgz_df,
  family = binomial(link = "logit"),
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000))
)

summary(model_extra, correlation = TRUE)

or_model_extra <- model_extra
ame_model_extra <- avg_slopes(model_extra, type = "response")
modelsummary(
  list("Odds Ratios" = or_model_extra, "Marginal Effects" = ame_model_extra),
  exponentiate = c(TRUE, FALSE),
  statistic = "conf.int",
  stars = TRUE,
  output = "gt"
)

anova(model_sign, model_extra)
check_collinearity(model_extra)
r2(model_extra)
```

```{r}
# MODEL COMPARISON TABLE: All models side by side

modelsummary(
  list(
    "Null"         = model_null,
    "For Pooling"  = model_for_pooling,
    "Full"         = model_full,
    "Significant"  = model_sign,
    "Extra"        = model_extra
  ),
  exponentiate = TRUE,
  statistic = "conf.int",
  stars = TRUE,
  gof_map = c("nobs", "logLik", "AIC", "BIC", "icc", "r2.marginal", "r2.conditional"),
  output = "gt"
)
```

```{r eval=F}
# Plots
# Coefficient plot (OR forest plot) for model_sign

coef_df <- tidy(model_sign, effects = "fixed", conf.int = TRUE, exponentiate = TRUE) %>%
  filter(term != "(Intercept)") %>%
  mutate(
    sig = case_when(
      p.value < 0.001 ~ "p < .001",
      p.value < 0.01  ~ "p < .01",
      p.value < 0.05  ~ "p < .05",
      TRUE            ~ "n.s."
    ),
    sig = factor(sig, levels = c("p < .001", "p < .01", "p < .05", "n.s."))
  )

p_coef <- ggplot(coef_df, aes(x = estimate, y = reorder(term, estimate), color = sig)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey50") +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  geom_point(size = 3) +
  scale_color_manual(
    values = c("p < .001" = "#d73027", "p < .01" = "#f46d43",
               "p < .05" = "#fdae61", "n.s." = "grey60")
  ) +
  labs(
    title    = "Odds Ratios — Significant Model",
    x        = "Odds Ratio (95% CI)",
    y        = NULL,
    color    = "Significance"
  ) +
  theme_bw(base_size = 12) +
  theme(legend.position = "bottom")

print(p_coef)


# Marginal effects plot (AME) 
ame_df <- as.data.frame(ame_model_sign) %>%
  filter(term != "(Intercept)") %>%
  mutate(
    sig = case_when(
      p.value < 0.001 ~ "p < .001",
      p.value < 0.01  ~ "p < .01",
      p.value < 0.05  ~ "p < .05",
      TRUE            ~ "n.s."
    ),
    sig = factor(sig, levels = c("p < .001", "p < .01", "p < .05", "n.s."))
  )

p_ame <- ggplot(ame_df, aes(x = estimate, y = reorder(term, estimate), color = sig)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  geom_point(size = 3) +
  scale_color_manual(
    values = c("p < .001" = "#d73027", "p < .01" = "#f46d43",
               "p < .05" = "#fdae61", "n.s." = "grey60")
  ) +
  labs(
    title = "Average Marginal Effects — Significant Model",
    x     = "AME on Pr(UB8=1)",
    y     = NULL,
    color = "Significance"
  ) +
  theme_bw(base_size = 12) +
  theme(legend.position = "bottom")

print(p_ame)


#  Predicted probabilities for key predictors
pred_ed5a <- ggpredict(model_sign, terms = "ED5A")
p_ed5a <- plot(pred_ed5a) +
  labs(title = "Predicted Probability by Mother's Education (ED5A)",
       x = "Education Level", y = "Pr(UB8 = 1)") +
  theme_bw(base_size = 12)

pred_windex <- ggpredict(model_sign, terms = "windex5")
p_windex <- plot(pred_windex) +
  labs(title = "Predicted Probability by Wealth Index",
       x = "Wealth Index (windex5)", y = "Pr(UB8 = 1)") +
  theme_bw(base_size = 12)

# For interaction in model_extra: ED5A * labor_women
pred_interaction <- ggpredict(model_extra, terms = c("ED5A", "labor_women"))
p_interaction <- plot(pred_interaction) +
  labs(title = "Interaction: Mother's Education × Women's Labour Force Participation",
       x = "Education Level", y = "Pr(UB8 = 1)", color = "Labor (women)") +
  theme_bw(base_size = 12)

print(p_ed5a)
print(p_windex)
print(p_interaction)


#  4. Random effects (regional variation) 
re_df <- ranef(model_sign, condVar = TRUE)$region_mics %>%
  tibble::rownames_to_column("region") %>%
  rename(intercept = `(Intercept)`) %>%
  mutate(
    se = sqrt(attr(ranef(model_sign, condVar = TRUE)$region_mics, "postVar")[1, 1, ]),
    lower = intercept - 1.96 * se,
    upper = intercept + 1.96 * se
  )

p_re <- ggplot(re_df, aes(x = intercept, y = reorder(region, intercept))) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0.3, color = "steelblue") +
  geom_point(size = 3, color = "steelblue") +
  labs(
    title = "Regional Random Intercepts (±1.96 SE)",
    x     = "Random Intercept (log-odds scale)",
    y     = "Region"
  ) +
  theme_bw(base_size = 12)

print(p_re)


#  5. Random slopes by region (model_extra: windex5 | region_mics) 
re_slopes_df <- ranef(model_extra, condVar = TRUE)$region_mics %>%
  tibble::rownames_to_column("region") %>%
  rename(intercept = `(Intercept)`, slope_windex = windex5)

p_slopes <- ggplot(re_slopes_df, aes(x = intercept, y = slope_windex, label = region)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  geom_point(size = 3, color = "steelblue") +
  ggrepel::geom_text_repel(size = 3) +
  labs(
    title = "Random Intercepts vs. Random Slopes for Wealth Index by Region",
    x     = "Random Intercept",
    y     = "Random Slope (windex5)"
  ) +
  theme_bw(base_size = 12)

print(p_slopes)


#  ICC and R² across models — summary bar chart 
model_list <- list(
  "Null"        = model_null,
  "For Pooling" = model_for_pooling,
  "Full"        = model_full,
  "Significant" = model_sign,
  "Extra"       = model_extra
)

r2_df <- purrr::map_dfr(names(model_list), function(nm) {
  r2_vals <- r2(model_list[[nm]])
  icc_val <- performance::icc(model_list[[nm]])$ICC_adjusted
  tibble::tibble(
    model        = nm,
    R2_marginal  = r2_vals$R2_marginal,
    R2_conditional = r2_vals$R2_conditional,
    ICC          = icc_val
  )
}) %>%
  tidyr::pivot_longer(-model, names_to = "metric", values_to = "value") %>%
  mutate(model = factor(model, levels = names(model_list)))

p_r2 <- ggplot(r2_df, aes(x = model, y = value, fill = metric)) +
  geom_col(position = "dodge") +
  scale_fill_manual(
    values = c("R2_marginal" = "#4575b4", "R2_conditional" = "#74add1", "ICC" = "#f46d43"),
    labels = c("ICC", "R² Conditional", "R² Marginal")
  ) +
  labs(
    title = "Model Fit Comparison: R² and ICC",
    x     = NULL,
    y     = "Value",
    fill  = NULL
  ) +
  ylim(0, 1) +
  theme_bw(base_size = 12) +
  theme(legend.position = "bottom")

print(p_r2)


# Combined layout for report 
(p_coef | p_ame) / (p_ed5a | p_windex) / p_r2 +
  plot_annotation(
    title   = "kyrgyzstan — Multilevel Model Results",
    caption = "MICS data; UB8 outcome; glmer with binomial logit link"
  )
```

# Model comparison

```{r}
# Splitting data

kgz_df <- kgz_df %>%
  mutate(strata_var = paste0(region_mics, "_", UB8))

train_idx <- createDataPartition(kgz_df$strata_var, p = 0.75, list = FALSE)
train_df  <- kgz_df[train_idx, ]
test_df   <- kgz_df[-train_idx, ]


# Refit all models on training data
# Swapping kgz_df to train_df in each model call
# Example:
model_null_train <- update(model_null, data = train_df)
model_pool_train <- update(model_for_pooling, data = train_df)
model_full_train <- update(model_full, data = train_df)
model_sign_train <- update(model_sign, data = train_df)
model_extra_train <- update(model_extra, data = train_df)

# Predict test set
# allow.new.levels = TRUE handles regions in test not seen in train
predict_prob <- function(model, newdata) {
  predict(model, newdata = newdata, type = "response", allow.new.levels = TRUE)
}

probs <- list(
  Null = predict_prob(model_null_train, test_df),
  For_Pooling = predict_prob(model_pool_train, test_df),
  Full = predict_prob(model_full_train, test_df),
  Significant = predict_prob(model_sign_train, test_df),
  Extra = predict_prob(model_extra_train, test_df)
)


# Metrics

metrics_df <- purrr::map2_dfr(
  probs, names(probs),
  ~ compute_metrics(.x, as.numeric(as.character(test_df$UB8)), .y)
)

# Comparison table

metrics_df %>%
  mutate(across(where(is.numeric), ~ round(.x, 3))) %>%
  gt::gt() %>%
  gt::tab_header(title = "Out-of-Sample Prediction Accuracy Comparison") %>%
  gt::fmt_number(columns = where(is.numeric), decimals = 3) %>%
  gt::cols_label(
    Model       = "Model",
    AUC         = "AUC",
    Accuracy    = "Accuracy",
    Sensitivity = "Sensitivity",
    Specificity = "Specificity",
    PPV         = "Precision (PPV)",
    F1          = "F1 Score",
    Brier       = "Brier Score",
    LogLoss     = "Log Loss"
  ) %>%
  gt::tab_style(
    style = gt::cell_fill(color = "#d4edda"),
    locations = gt::cells_body(
      rows = AUC == max(AUC)
    )
  )

# ROC curves plot (all models)
roc_list <- purrr::map2(probs, names(probs), function(p, nm) {
  roc_obj <- roc(as.numeric(as.character(test_df$UB8)), p, quiet = TRUE)
  data.frame(
    model       = paste0(nm, " (AUC=", round(auc(roc_obj), 3), ")"),
    sensitivity = roc_obj$sensitivities,
    specificity = roc_obj$specificities
  )
})

roc_df <- dplyr::bind_rows(roc_list)

p_roc <- ggplot(roc_df, aes(x = 1 - specificity, y = sensitivity, color = model)) +
  geom_line(linewidth = 0.9) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50") +
  scale_color_brewer(palette = "Set1") +
  labs(
    title = "ROC Curves — Out-of-Sample Performance",
    x     = "1 - Specificity (FPR)",
    y     = "Sensitivity (TPR)",
    color = NULL
  ) +
  theme_bw(base_size = 12) +
  theme(legend.position = "bottom")

print(p_roc)


# Calibration plot — predicted vs. observed probabilities
cal_df <- purrr::map2_dfr(probs, names(probs), function(p, nm) {
  data.frame(
    model    = nm,
    pred     = p,
    observed = as.numeric(as.character(test_df$UB8))
  )
})

# Bin predictions into deciles per model
cal_binned <- cal_df %>%
  dplyr::group_by(model) %>%
  dplyr::mutate(bin = ntile(pred, 10)) %>%
  dplyr::group_by(model, bin) %>%
  dplyr::summarise(
    mean_pred = mean(pred),
    mean_obs  = mean(observed),
    .groups   = "drop"
  )

p_cal <- ggplot(cal_binned, aes(x = mean_pred, y = mean_obs, color = model)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50") +
  geom_line(linewidth = 0.7, alpha = 0.8) +
  geom_point(size = 2) +
  scale_color_brewer(palette = "Set1") +
  labs(
    title = "Calibration Plot — Predicted vs. Observed Probabilities",
    x     = "Mean Predicted Probability",
    y     = "Mean Observed Proportion",
    color = NULL
  ) +
  theme_bw(base_size = 12) +
  theme(legend.position = "bottom")

print(p_cal)
```