---
title: "Pooled model for all countries"
author: "Berdiyeva Sofiya"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: true
    highlight: tango
---

# Pipeline

We ran and assessed the performance of one pooled model.

# Key takeaways

- Both country and region-level groupings are justified (ICC>0.05)

- Significant predictors: child's age in months [+], caregiver's education [+], number of books in the household [+], having more than 2 playthings [-], child being developmentally on track [+], location in rural area [-], household size [-], household wealth index quintile [+], car presence [-].

- Kazakhstan has the highest intercept, which is consistent with the highest enrollment rates in the region.

- Model accuracy is 0.712.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache=T)
htmlwidgets::setWidgetIdSeed(123)

source(here::here("R", "00_load_packages.R"))
source(here::here("R", "01_data_cleaning.R"))
source(here::here("R", "02_descriptives.R"))
set.seed(42)
```

# Loading country data

## Kazakhstan

```{r}
# Indiv-level data
kazakhstan_mics <- get_and_preprocess_mics_data("Kazakhstan")

# Region-level data
region_kz <- read_csv("../data/raw/regional_data_Kazakhstan_2024.csv")
```

### Preprocessing individual level data 

#### Dimensionality reduction

```{r}
kazakhstan_mics <- kazakhstan_mics %>% 
  
  # Learning materials
  recode_playthings(books_var = "EC1",
                    playthings_vars = c("EC2A", "EC2B", "EC2C")) %>% 
  
  # Inadequate supervision
  recode_supervision(supervision_vars = c("EC3A", "EC3B")) %>% 
  
  # Child discipline - psychological aggression
  recode_psy_aggression(var1 = "UCD2D", var2 = "UCD2H") %>% 
  
  # Early stimulation and responsive care
  recode_early_stimulation(stimulation_vars = c("EC5A", "EC5B", "EC5C",
                                                "EC5D", "EC5E", "EC5F")) %>% 
  
  # ECDI
  recode_ecdi_and_track(age_col = "CAGE",
                        ec_regular = paste0("EC", 6:23),
                        ec_frequency = "EC24",
                        ec_comparison = "EC25") %>% 
  mutate(UB8=if_else(UB8==1, 1,0)) %>% 
  mutate(HC10E=if_else(HC10E==1, 1,0)) %>% 

  # Caregiver's education
  recode_education_scheme("ED5A", "KZ") %>% 
  
  # Sample one child from household
  sample_one_per_group(group_var = "house_index",
                       seed = 42) %>% 

  # Omit rows with missing values for key predictors
  filter_at(vars(ED5A, HL13, HL17), all_vars(!is.na(.)))
```

#### Imputation

```{r}
kazakhstan_mics$UB8[is.na(kazakhstan_mics$UB8)] <- 0
```

### Combining data for multilevel modeling

```{r}
# If we decide to grand-center anything
region_kz <- region_kz %>%
  # Log-transform to compress the scale
  mutate(places_in_pre_total = places_in_pre_total/1000,
         popul_total_01012024 = popul_total_01012024/1000000,
         pre_teachers_total_2024 = pre_teachers_total_2024/100)
region_kz <- grand_mean_centering(region_kz, colnames(region_kz)[-c(1:3)])
region_kz <- region_kz %>%
  # Log-transform to compress the scale
  mutate(popul_density_log = log(popul_density),
         total_pre_2024_log = log(total_pre_2024))

kaz_df <- kazakhstan_mics %>%
  mutate(region_mics = as.character(as_factor(HH7))) %>% 
  left_join(region_kz,
            by = "region_mics",
            keep = FALSE)
```

## Kyrgyzstan

```{r}
# Indiv-level data
kyrgyzstan_mics <- get_and_preprocess_mics_data("Kyrgyzstan")

# Region-level data
region_kg <- read_csv("../data/raw/regional_data_Kyrgyzstan_2023.csv")
```

### Preprocessing individual level data 

#### Dimensionality reduction

```{r}
kyrgyzstan_mics <- kyrgyzstan_mics %>% 
  
  # Learning materials
  recode_playthings(books_var = "EC1",
                    playthings_vars = c("EC2A", "EC2B", "EC2C")) %>% 
  
  # Inadequate supervision
  recode_supervision(supervision_vars = c("EC3A", "EC3B")) %>% 
  
  # Child discipline - psychological aggression
  recode_psy_aggression(var1 = "UCD2D", var2 = "UCD2H") %>% 
  
  # Early stimulation and responsive care
  recode_early_stimulation_v2(
    activity_prefixes = c("EC5A", "EC5B", "EC5C", "EC5D", "EC5E", "EC5F")) %>% 
  
  # ECDI
  recode_ecdi_and_track(age_col = "CAGE",
                        ec_regular = paste0("EC", 21:38),
                        ec_frequency = "EC39",
                        ec_comparison = "EC40") %>% 
  mutate(UB8=if_else(UB8==1, 1,0)) %>% 
  mutate(HC10E=if_else(HC10E==1, 1,0)) %>% 
  
  # Social transfers
  recode_economic_assistance(unit_base = "ST4U$",
                                          number_base = "ST4N$",
                                          n_programs = 5,
                                          months_threshold = 3) %>% 
  # Caregiver's education
  recode_education_scheme("ED5A", "KG") %>% 
  
  # Exclude children under 3
  filter(CAGE>35) %>% 
  
  # Sample one child from household
  sample_one_per_group(group_var = "house_index",
                       seed = 42) %>% 

  # Omit rows with missing values for key predictors
  filter_at(vars(ED5A, HL13, HL17), all_vars(!is.na(.)))
```

#### Imputation

```{r}
kyrgyzstan_mics$UB8[is.na(kyrgyzstan_mics$UB8)] <- 0
```

### Combining data for multilevel modeling

```{r}
# If we decide to grand-center anything
region_kg <- region_kg %>%
  # Log-transform to compress the scale
  mutate(popul_mln_total_01012023 = popul_thsnd_total_01012023/1000)
region_kg <- grand_mean_centering(region_kg, colnames(region_kg)[-c(1:2)])
region_kg <- region_kg %>%
  # Log-transform to compress the scale
  mutate(popul_density_log = log(popul_density),
         total_pre_2023_log = log(total_pre_2023),
         expen_on_pre_2023_log = log(expen_on_pre_2023))

kgz_df <- kyrgyzstan_mics %>%
  mutate(region_mics = as.character(as_factor(HH7))) %>% 
  left_join(region_kg,
            by = "region_mics",
            keep = FALSE)
```

## Turkmenistan

```{r}
# Indiv-level data
turkmenistan_mics <- get_and_preprocess_mics_data("Turkmenistan")

# Region-level data
region_tm <- read_csv("../data/raw/regional_data_Turkmenistan_2024.csv")
```

### Preprocessing individual level data 

#### Dimensionality reduction

```{r}
turkmenistan_mics <- turkmenistan_mics %>% 
  
  # Learning materials
  recode_playthings(books_var = "EC1",
                    playthings_vars = c("EC2A", "EC2B", "EC2C")) %>% 
  
  # Inadequate supervision
  recode_supervision(supervision_vars = c("EC3A", "EC3B")) %>% 
  
  # Child discipline - psychological aggression
  recode_psy_aggression(var1 = "UCD2D", var2 = "UCD2H") %>% 
  
  # Early stimulation and responsive care
  recode_early_stimulation(stimulation_vars = c("EC5A", "EC5B", "EC5C",
                                                "EC5D", "EC5E", "EC5F")) %>% 
  
  # ECDI
  recode_ecdi_and_track(age_col = "CAGE",
                        ec_regular = paste0("EC", 6:23),
                        ec_frequency = "EC24",
                        ec_comparison = "EC25") %>% 
  mutate(UB8=if_else(UB8==1, 1,0)) %>% 
  mutate(HC10E=if_else(HC10E==1, 1,0)) %>% 
  
  # Social transfers
  recode_economic_assistance(unit_base = "ST6U$",
                                          number_base = "ST6N$",
                                          n_programs = 5,
                                          months_threshold = 3) %>% 
  # Caregiver's education
  recode_education_scheme("ED5A", "TM") %>% 
  
  # Exclude children under 3
  filter(CAGE>35) %>% 
  
  # Sample one child from household
  sample_one_per_group(group_var = "house_index",
                       seed = 42) %>% 

  # Omit rows with missing values for key predictors
  filter_at(vars(ED5A, HL13, HL17), all_vars(!is.na(.)))
```

#### Imputation

```{r}
turkmenistan_mics$UB8[is.na(turkmenistan_mics$UB8)] <- 0
```

### Combining data for multilevel modeling

```{r}
# If we decide to grand-center anything
region_tm <- region_tm %>%
  # Log-transform to compress the scale
  mutate(popul_mln_total_01012025 = popul_thsnd_total_01012025/1000)
region_tm <- grand_mean_centering(region_tm, colnames(region_tm)[-c(1:2)])
region_tm <- region_tm %>%
  # Log-transform to compress the scale
  mutate(popul_density_log = log(popul_density),
         total_pre_end_2024_log = log(total_pre_end_2024))

tkm_df <- turkmenistan_mics %>%
  mutate(region_mics = as.character(as_factor(HH7))) %>% 
  left_join(region_tm,
            by = "region_mics",
            keep = FALSE)
```

## Uzbekistan

```{r}
# Indiv-level data
uzbekistan_mics <- get_and_preprocess_mics_data("Uzbekistan")

# Region-level data
region_uz <- read_csv("../data/raw/regional_data_Uzbekistan_2022.csv")
```

### Preprocessing individual level data 

#### Dimensionality reduction

```{r}
uzbekistan_mics <- uzbekistan_mics %>% 
  
  # Learning materials
  recode_playthings(books_var = "EC1",
                    playthings_vars = c("EC2A", "EC2B", "EC2C")) %>% 
  
  # Inadequate supervision
  recode_supervision(supervision_vars = c("EC3A", "EC3B")) %>% 
  
  # Child discipline - psychological aggression
  recode_psy_aggression(var1 = "UCD2D", var2 = "UCD2H") %>% 
  
  # Early stimulation and responsive care
  recode_early_stimulation_v2(
    activity_prefixes = c("EC5A", "EC5B", "EC5C", "EC5D", "EC5E", "EC5F")) %>% 
  
  # ECDI
  recode_ecdi_and_track(age_col = "CAGE",
                        ec_regular = paste0("EC", 21:38),
                        ec_frequency = "EC39",
                        ec_comparison = "EC40") %>% 
  mutate(UB8=if_else(UB8==1, 1,0)) %>% 
  mutate(HC10E=if_else(HC10E==1, 1,0)) %>% 
  
  # Social transfers
  recode_economic_assistance(unit_base = "ST4U$",
                                          number_base = "ST4N$",
                                          n_programs = 5,
                                          months_threshold = 3) %>% 
  # Caregiver's education
  recode_education_scheme("ED5A", "UZ") %>% 
  
  # Exclude children under 3
  filter(CAGE>35) %>% 
  
  # Sample one child from household
  sample_one_per_group(group_var = "house_index",
                       seed = 42) %>% 

  # Omit rows with missing values for key predictors
  filter_at(vars(ED5A, HL13, HL17), all_vars(!is.na(.)))
```

#### Imputation

```{r}
uzbekistan_mics$UB8[is.na(uzbekistan_mics$UB8)] <- 0
```

### Combining data for multilevel modeling

```{r}
# If we decide to grand-center anything
region_uz <- region_uz %>%
  # Log-transform to compress the scale
  mutate(popul_mln_total_01012022 = popul_thsnd_total_01012022/1000)
region_uz <- grand_mean_centering(region_uz, colnames(region_uz)[-c(1:2)])
region_uz <- region_uz %>%
  # Log-transform to compress the scale
  mutate(popul_density_2022_log = log(popul_density_2022),
         total_pre_2022_log = log(total_pre_2022),
         places_in_pre_total_2022_log = log(places_in_pre_total_2022),
         pre_teachers_total_2022_log = log(pre_teachers_total_2022))

uzb_df <- uzbekistan_mics %>%
  mutate(region_mics = as.character(as_factor(HH7))) %>% 
  left_join(region_uz,
            by = "region_mics",
            keep = FALSE)
```

# Combining all dataframes

```{r}
# Unifying variable names
tkm_df <- tkm_df %>% rename(popul_mln_total = popul_mln_total_01012025,
                            total_pre = total_pre_end_2024)
uzb_df <- uzb_df %>% rename(popul_mln_total = popul_mln_total_01012022,
                            total_pre = total_pre_2022,
                            popul_density = popul_density_2022)
uzb_df <- uzb_df %>%
  rename_with(~paste0("EC", as.numeric(gsub("EC", "", .)) - 15),
              all_of(paste0("EC", 21:40)))
kgz_df <- kgz_df %>% rename(popul_mln_total = popul_mln_total_01012023,
                            total_pre = total_pre_2023)
kgz_df <- kgz_df %>%
  rename_with(~paste0("EC", as.numeric(gsub("EC", "", .)) - 15),
              all_of(paste0("EC", 21:40)))

kaz_df <- kaz_df %>% rename(total_pre = total_pre_2024,
                            popul_mln_total = popul_total_01012024)

```


```{r}
datframes <- list(
  list(df = uzb_df, country = "Uzbekistan", code = "uz"),
  list(df = kaz_df, country = "Kazakhstan", code = "kz"),
  list(df = kgz_df, country = "Kyrgyzstan", code = "kg"),
  list(df = tkm_df, country = "Turkmenistan", code = "tm")
)

needed_cols <- c("UB8", "CAGE", "ED5A", "EC1", "two_plus_playthings", "supervised_inadequately", "psy_aggression", "early_stimulation", "dev_on_track", "HH6", "HH48", "HL13", "HL17", "windex5", "region", "popul_mln_total", "total_pre", "popul_density", "HC10E")

combined_df <- datframes %>%
  map_df(function(x) {
    x$df %>%
      select(any_of(needed_cols)) %>%  
      mutate(country = x$country, .before = 1)
  }) %>% 
  mutate(total_pre_log = log(total_pre),
         popul_density_log = log(popul_density))
```

```{r}
# THREE-LEVEL MODEL
# Level 1: Individual (child)
# Level 2: Region within country
# Level 3: Country

# ---- Null model ----
model_null_3l <- glmer(
  UB8 ~ 1 + (1 | country/region),
  data = combined_df,
  family = binomial(link = "logit"),
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000))
)

vpc_null_3l <- performance::icc(model_null_3l, by_group = TRUE)
print(vpc_null_3l)
# output tells the three-level structure is justified:
# Country level (L3): ICC = 0.058 — about 6% of variance in UB8 is attributable to country-level differences. Modest but meaningful, justifying the third level.
# Region level (L2): ICC = 0.099 — about 10% of variance is attributable to region-level clustering (this is the combined region+country variance). So regions matter more than countries for explaining variation in UB8.
```

# Main model

```{r}
model_3l <- glmer(
UB8 ~ 
    # Level 1 predictors (child/household)
    CAGE + # Age in months
    ED5A + # mother's/caregiver's education
    EC1 + #Three or more books
    two_plus_playthings +
    supervised_inadequately +
    psy_aggression +
    early_stimulation +
    dev_on_track +
    HH6 + # Rural/urban
    HH48 + # household size
    HL13 + # Natural mother lives in the household
    HL17 + # Natural father lives in the household
    windex5 +
    HC10E + # Family has a car
  
      # Level 2 predictors (structure)
    popul_mln_total +
    popul_density_log +
    total_pre_log +
    # Random structure
    (1 | country/region),
  
  data = combined_df,
  family = binomial(link = "logit"),
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000))
)

summary(model_3l, correlation = TRUE)

or_model_3l  <- model_3l
ame_model_3l <- avg_slopes(model_3l, type = "response")

modelsummary(
  list("Odds Ratios" = or_model_3l, "Marginal Effects" = ame_model_3l),
  exponentiate = c(TRUE, FALSE),
  statistic = "conf.int",
  stars = TRUE,
  output = "gt"
)

anova(model_null_3l, model_3l)
check_collinearity(model_3l)
r2(model_3l)
```


# Model assessment
```{r}
# ICC decomposition across levels
icc_3l <- performance::icc(model_3l, by_group = TRUE)
print(icc_3l)

# Variance partition: how much variance at each level
vc <- as.data.frame(VarCorr(model_3l))
vc %>%
  select(grp, var1, vcov, sdcor) %>%
  mutate(
    pct_variance = round(vcov / sum(vcov) * 100, 1)
  ) %>%
  print()
```

# Visualizations

```{r}
# 1. Country-level random effects
re_country <- ranef(model_3l, condVar = TRUE)$country %>%
  tibble::rownames_to_column("country") %>%
  rename(intercept = `(Intercept)`) %>%
  mutate(
    se    = sqrt(attr(ranef(model_3l, condVar = TRUE)$country, "postVar")[1, 1, ]),
    lower = intercept - 1.96 * se,
    upper = intercept + 1.96 * se
  )

p_re_country <- ggplot(re_country, aes(x = intercept, y = reorder(country, intercept))) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0.2, color = "#e41a1c") +
  geom_point(size = 3, color = "#e41a1c") +
  labs(
    title = "Country-Level Random Intercepts (±1.96 SE)",
    x     = "Random Intercept (log-odds scale)",
    y     = NULL
  ) +
  theme_bw(base_size = 12)


# ---- 2. Region-level random effects, faceted by country ----
re_region <- ranef(model_3l, condVar = TRUE)$`region:country` %>%
  tibble::rownames_to_column("region_country") %>%
  rename(intercept = `(Intercept)`) %>%
  tidyr::separate(region_country, into = c("region", "country"), sep = ":") %>%
  mutate(
    se    = sqrt(attr(ranef(model_3l, condVar = TRUE)$`region:country`, "postVar")[1, 1, ]),
    lower = intercept - 1.96 * se,
    upper = intercept + 1.96 * se
  )

p_re_region <- ggplot(re_region, aes(x = intercept, y = reorder(region, intercept))) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  geom_errorbar(aes(xmin = lower, xmax = upper), 
              height = 0.3, color = "steelblue", alpha = 0.6,
              orientation = "y") +
  geom_point(size = 2, color = "steelblue") +
  facet_wrap(~ country, scales = "free_y", ncol = 2) +
  labs(
    title = "Region-Level Random Intercepts by Country (±1.96 SE)",
    x     = "Random Intercept (log-odds scale)",
    y     = NULL
  ) +
  theme_bw(base_size = 11) +
  theme(strip.background = element_rect(fill = "#f0f0f0"))


# ---- 3. Variance partition bar chart ----
vc_plot <- vc %>%
  select(grp, vcov) %>%
  mutate(
    grp  = recode(grp,
      "country"             = "Country (L3)",
      "country:region" = "Region (L2)",
      "Residual"            = "Individual (L1)"
    ),
    pct = vcov / sum(vcov) * 100
  )

p_vpc <- ggplot(vc_plot, aes(x = "", y = pct, fill = grp)) +
  geom_col(width = 0.4) +
  geom_text(aes(label = paste0(round(pct, 1), "%")),
            position = position_stack(vjust = 0.5), size = 4) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Variance Partition by Level",
    x     = NULL,
    y     = "% of Total Variance",
    fill  = NULL
  ) +
  theme_bw(base_size = 12) +
  theme(legend.position = "bottom")


p_re_country
p_vpc
p_re_region

```

# Assessing accuracy

```{r}
set.seed(42)

combined_df <- combined_df %>%
  mutate(strata_var = paste0(country, "_", region, "_", UB8))

train_idx_3l <- createDataPartition(combined_df$strata_var, p = 0.75, list = FALSE)
train_3l     <- combined_df[train_idx_3l, ]
test_3l      <- combined_df[-train_idx_3l, ]

model_3l_train <- update(model_3l, data = train_3l)

probs_3l <- predict(model_3l_train, newdata = test_3l,
                    type = "response", allow.new.levels = TRUE)

metrics_3l <- compute_metrics(
  probs   = probs_3l,
  actual  = as.numeric(as.character(test_3l$UB8)),
  model_name = "Three-Level"
)

kable(metrics_3l)
```

