---
title: "Uzbekistan MICS 2021-2022"
author: "Berdiyeva Sofiya"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: true
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache=T, warning = F)
htmlwidgets::setWidgetIdSeed(123)

source(here::here("R", "00_load_packages.R"))
source(here::here("R", "01_data_cleaning.R"))
source(here::here("R", "02_descriptives.R"))
set.seed(42)
```

# What a reader might need to know about the modeling process

## Pipeline

The development of models and variable selection was following the logic below:

1. First, we run a null model to check whether multilevel approach is justified at all

2. Next, we run a model with the most exhaustive list of predictors that are shared accross countries' datasets, called a "model for pooling". This model is run through anova to see whether it fits the data significantly better than the null model.

3. After that, we do a little robustness check, running (almost) the same model for pooling through the regular logistic regression rather than multilevel one.

4. Then, we include all variables available for the country (that are of our interest) into a "full" model and check whether it is significantly different in prediction than a model for pooling.

5. Also, we are tracking r^2 and collinearity metrics all way along.

6. As a next step, we remove all non-significant variables from a full model, calling this model "model_sign", and based on it, conduct some questionable experiments with interactions and random slopes and call it 'model_extra'

7. Finally, we are actually doing some machine learning-style performance evaluation of all of the models, preserving the formulas, but re-training them on randomly (but within stratas) split  data.

## Key takeaways for Uzbekistan

- Multilevel modeling is justified, since ICC perfectly passed 0.001 point over a commonly used threshold (ICC=0.051).

- Robustness check shows unstable results for having more than two playthings, psychological aggression, father living in the household.

- Significant model won in performance by accuracy equal to 0.624 (the lowest result across countries), see further [metrics](#performance-comparison-table)), and included the variables: child's age in months [+], caregiver's education [+], number of books in the household [+], having more than two playthings [-], car presence [+]. None of the structural predictors were significant.

- Highest conditional R-squared reached 0.169 (See a model comparison [table](#model-comparison-table))


# Loading data

```{r}
# Indiv-level data
uzbekistan_mics <- get_and_preprocess_mics_data("Uzbekistan")

# Region-level data
region_uz <- read_csv("../data/raw/regional_data_Uzbekistan_2022.csv")
```

# Preprocessing individual level data 

## Dimensionality reduction

```{r}
uzbekistan_mics <- uzbekistan_mics %>% 
  
  # Learning materials
  recode_playthings(books_var = "EC1",
                    playthings_vars = c("EC2A", "EC2B", "EC2C")) %>% 
  
  # Inadequate supervision
  recode_supervision(supervision_vars = c("EC3A", "EC3B")) %>% 
  
  # Child discipline - psychological aggression
  recode_psy_aggression(var1 = "UCD2D", var2 = "UCD2H") %>% 
  
  # Early stimulation and responsive care
  recode_early_stimulation_v2(
    activity_prefixes = c("EC5A", "EC5B", "EC5C", "EC5D", "EC5E", "EC5F")) %>% 
  
  # ECDI
  recode_ecdi_and_track(age_col = "CAGE",
                        ec_regular = paste0("EC", 21:38),
                        ec_frequency = "EC39",
                        ec_comparison = "EC40") %>% 
  mutate(UB8=if_else(UB8==1, 1,0)) %>% 
  mutate(HC10E=if_else(HC10E==1, 1,0)) %>% 

  
  # Social transfers
  recode_economic_assistance(unit_base = "ST4U$",
                                          number_base = "ST4N$",
                                          n_programs = 5,
                                          months_threshold = 3) %>% 
  # Caregiver's education
  recode_education_scheme("ED5A", "UZ") %>% 
  
  # Exclude children under 3
  filter(CAGE>35) %>% 
  
  # Sample one child from household
  sample_one_per_group(group_var = "house_index",
                       seed = 42) %>% 

  # Omit rows with missing values for key predictors
  filter_at(vars(ED5A, HL13, HL17), all_vars(!is.na(.)))
```

## Imputation

```{r}
uzbekistan_mics$UB8[is.na(uzbekistan_mics$UB8)] <- 0
```


# Correlations and associations

```{r, warning=F}

results <- pairwise_chi_cramers_parallel(
  uzbekistan_mics[, names(uzbekistan_mics) %in% c("UB2", "ED5A","UB8","HH48", "HH51","HC10E", "HH7", "HH6", "HL13", "HL17", "windex5", "ED4", "playthings_sum", "two_plus_playthings", "supervised_inadequately", "psy_aggression", "early_stimulation", "ecdi_sum", "dev_on_track", "received_econ_assist" )], n_cores = 4, p_threshold = 0.05, use_bonferroni = FALSE)


# Get summary
summary_stats(results)

# Create heatmap
plot <- plot_cramers_heatmap(results, text_size_x = 11, text_size_y = 11)
print(plot)

# View top associations
top_20 <- get_top_associations(results, n = 20)
print(top_20)
```

# Combining data for multilevel modeling

```{r}
# If we decide to grand-center anything
region_uz <- region_uz %>%
  # Log-transform to compress the scale
  mutate(popul_mln_total_01012022 = popul_thsnd_total_01012022/1000)
region_uz <- grand_mean_centering(region_uz, colnames(region_uz)[-c(1:2)])
region_uz <- region_uz %>%
  # Log-transform to compress the scale
  mutate(popul_density_2022_log = log(popul_density_2022),
         total_pre_2022_log = log(total_pre_2022),
         places_in_pre_total_2022_log = log(places_in_pre_total_2022),
         pre_teachers_total_2022_log = log(pre_teachers_total_2022))

uzb_df <- uzbekistan_mics %>%
  mutate(region_mics = as.character(as_factor(HH7))) %>% 
  left_join(region_uz,
            by = "region_mics",
            keep = FALSE)
```

# Modeling

## Baseline models

### Null model

```{r}
# RANDOM INTERCEPT WITH FIXED SLOPES

# Null model (unconditional model) - to calculate ICC
model_null <- glmer(
  UB8 ~ 1 + (1 | region_mics),
  data = uzb_df,
  family = binomial(link = "logit"),
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000))
)

# Calculate Intraclass Correlation Coefficient (ICC)
# ICC tells us how much variance is due to clustering
# An ICC > 0.05 or 0.10 or typically justifies using a multilevel model
vpc_null <- performance::icc(model_null)
print(paste("ICC:", round(vpc_null$ICC_adjusted, 3)))
```

### All individual predictors + universal regional predictors

```{r}
# Baseline random intercept model
model_for_pooling <- glmer(
  UB8 ~ 
    # Level 1 predictors (child/household)
    CAGE + # Age in months
    ED5A + # mother's/caregiver's education
    EC1 + #Three or more books
    two_plus_playthings +
    supervised_inadequately +
    psy_aggression +
    early_stimulation +
    dev_on_track +
    HH6 + # Rural/urban
    HH48 + # household size
    HL13 + # Natural mother lives in the household
    HL17 + # Natural father lives in the household
    windex5 +
    HC10E + # Family has a car

    
    # Level 2 predictors (structure)
    popul_mln_total_01012022 +
    popul_density_2022_log +
    total_pre_2022_log +

    # Random intercept for region
    (1 | region),
  
  data = uzb_df,
  family = binomial(link = "logit"),
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000))
)

# Model summary
summary(model_for_pooling, correlation =TRUE)

or_model <- model_for_pooling
ame_model <- avg_slopes(model_for_pooling, type = "response")

modelsummary(
  list(
    "Odds Ratios" = or_model,
    "Marginal Effects" = ame_model
  ),
  exponentiate = c(TRUE, FALSE),
  statistic = "conf.int",
  stars = TRUE,
  output = "gt"
)
```

```{r}
anova(model_null, model_for_pooling)
check_collinearity(model_for_pooling)
r2(model_for_pooling)
r2(model_null)
```

```{r}
# Robustness check
model_fe_region <- glm(
  UB8 ~ CAGE + # Age in months
    ED5A + # mother's/caregiver's education
    EC1 + #Three or more books
    two_plus_playthings +
    supervised_inadequately +
    psy_aggression +
    early_stimulation +
    dev_on_track +
    HH6 + # Rural/urban
    HH48 + # household size
    HL13 + # Natural mother lives in the household
    HL17 + # Natural father lives in the household
    windex5 +
    HC10E + # Family has a car
    
    # Level 2 predictors (structure)
    popul_mln_total_01012022 +
    popul_density_2022_log +
    total_pre_2022_log +
    factor(region),
  data = uzb_df,
  family = binomial
)

modelsummary(
  list(
    "Multilevel model" = model_for_pooling,
    "Region Fixed Effects Model" = model_fe_region
  ),
  exponentiate = TRUE
)
```

### All available country-specific variables, fixed slopes

```{r}
model_full <- glmer(
    UB8 ~ 
    # Level 1 predictors (child/household)
    CAGE + # Age in months
    ED5A + # mother's/caregiver's education
    EC1 + #Three or more books
    two_plus_playthings +
    supervised_inadequately +
    psy_aggression +
    early_stimulation +
    dev_on_track +
    received_econ_assist + # Social transfers!
    HH6 + # Rural/urban
    HH48 + # household size
    HL13 + # Natural mother lives in the household
    HL17 + # Natural father lives in the household
    windex5 +
    HC10E + # Family has a car
    
    # Level 2 predictors (structure)
    popul_mln_total_01012022 +
    popul_density_2022_log +
    total_pre_2022_log +   
    places_in_pre_total_2022_log +
    pre_teachers_total_2022_log +
    labor_force_part_men_2022_c + 
    labor_force_part_women_2022_c +
    private_pre_share_c +
    
    # Random intercept for region
    (1 | region),
  
  data = uzb_df,
  family = binomial(link = "logit"),
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000))
)

summary(model_full, correlation = TRUE)

or_model_full <- model_full
ame_model_full <- avg_slopes(model_full, type = "response")
modelsummary(
  list("Odds Ratios" = or_model_full, "Marginal Effects" = ame_model_full),
  exponentiate = c(TRUE, FALSE),
  statistic = "conf.int",
  stars = TRUE,
  output = "gt"
)

anova(model_for_pooling, model_full)
check_collinearity(model_full)
r2(model_full)
```


## Reduced models
### Only significant fixed effects (p < .05) from model_full

```{r}
model_sign <- glmer(
    UB8 ~ 
    # Level 1 predictors (child/household)
    CAGE + # Age in months
    ED5A + # mother's/caregiver's education
    EC1 + #Three or more books
    two_plus_playthings +
    HC10E + # Family has a car
    
    # Random intercept for region
    (1 | region),  
  data = uzb_df,
  family = binomial(link = "logit"),
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000))
)

summary(model_sign, correlation = TRUE)

or_model_sign <- model_sign
ame_model_sign <- avg_slopes(model_sign, type = "response")
modelsummary(
  list("Odds Ratios" = or_model_sign, "Marginal Effects" = ame_model_sign),
  exponentiate = c(TRUE, FALSE),
  statistic = "conf.int",
  stars = TRUE,
  output = "gt"
)

anova(model_full, model_sign)
check_collinearity(model_sign)
r2(model_sign)
```

### Extra model: interactions + random slopes

```{r}
model_extra <- glmer(
  UB8 ~ 
    # Level 1 predictors (child/household)
    CAGE + # Age in months
    ED5A + # mother's/caregiver's education
    EC1 + #Three or more books
    windex5 +
    two_plus_playthings+
    HC10E + # Family has a car
    
    # Level 2 predictors (structure)
    popul_density_2022_log +
    total_pre_2022_log +
    places_in_pre_total_2022_log +
    
    # Random slopes
    (1 + windex5 | region_mics),
  
  data = uzb_df,
  family = binomial(link = "logit"),
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000))
)

summary(model_extra, correlation = TRUE)

or_model_extra <- model_extra
ame_model_extra <- avg_slopes(model_extra, type = "response")
modelsummary(
  list("Odds Ratios" = or_model_extra, "Marginal Effects" = ame_model_extra),
  exponentiate = c(TRUE, FALSE),
  statistic = "conf.int",
  stars = TRUE,
  output = "gt"
)

anova(model_sign, model_extra)
check_collinearity(model_extra)
r2(model_extra)
```

# Model comparison table

```{r}
modelsummary(
  list(
    "Null"         = model_null,
    "For Pooling"  = model_for_pooling,
    "Full"         = model_full,
    "Significant"  = model_sign,
    "Extra"        = model_extra
  ),
  exponentiate = TRUE,
  statistic = "conf.int",
  stars = TRUE,
  gof_map = c("nobs", "logLik", "AIC", "BIC", "icc", "r2.marginal", "r2.conditional"),
  output = "gt"
)
```

```{r}
# Plots
# Coefficient plot (OR forest plot) for model_sign

coef_df <- tidy(model_sign, effects = "fixed", conf.int = TRUE, exponentiate = TRUE) %>%
  filter(term != "(Intercept)") %>%
  mutate(
    sig = case_when(
      p.value < 0.001 ~ "p < .001",
      p.value < 0.01  ~ "p < .01",
      p.value < 0.05  ~ "p < .05",
      TRUE            ~ "n.s."
    ),
    sig = factor(sig, levels = c("p < .001", "p < .01", "p < .05", "n.s."))
  )

p_coef <- ggplot(coef_df, aes(x = estimate, y = reorder(term, estimate), color = sig)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey50") +
  geom_errorbar(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  geom_point(size = 3) +
  scale_color_manual(
    values = c("p < .001" = "#d73027", "p < .01" = "#f46d43",
               "p < .05" = "#fdae61", "n.s." = "grey60")
  ) +
  labs(
    title    = "Odds Ratios — Significant Model",
    x        = "Odds Ratio (95% CI)",
    y        = NULL,
    color    = "Significance"
  ) +
  theme_bw(base_size = 12) +
  theme(legend.position = "bottom")

print(p_coef)


# Marginal effects plot (AME) 
ame_df <- as.data.frame(ame_model_sign) %>%
  filter(term != "(Intercept)") %>%
  mutate(
    sig = case_when(
      p.value < 0.001 ~ "p < .001",
      p.value < 0.01  ~ "p < .01",
      p.value < 0.05  ~ "p < .05",
      TRUE            ~ "n.s."
    ),
    sig = factor(sig, levels = c("p < .001", "p < .01", "p < .05", "n.s."))
  )

p_ame <- ggplot(ame_df, aes(x = estimate, y = reorder(term, estimate), color = sig)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  geom_point(size = 3) +
  scale_color_manual(
    values = c("p < .001" = "#d73027", "p < .01" = "#f46d43",
               "p < .05" = "#fdae61", "n.s." = "grey60")
  ) +
  labs(
    title = "Average Marginal Effects — Significant Model",
    x     = "AME on Pr(UB8=1)",
    y     = NULL,
    color = "Significance"
  ) +
  theme_bw(base_size = 12) +
  theme(legend.position = "bottom")

print(p_ame)


#  Predicted probabilities for key predictors
pred_ed5a <- ggpredict(model_sign, terms = "ED5A [all]", bias_correction = TRUE)
p_ed5a <- plot(pred_ed5a) +
  labs(title = "Predicted Probability by Mother's Education (ED5A)",
       x = "Education Level", y = "Pr(UB8 = 1)") +
  theme_bw(base_size = 12)+
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

pred_windex <- ggpredict(model_sign, terms = "HC10E [all]", bias_correction = TRUE)
p_windex <- plot(pred_windex) +
  labs(title = "Predicted Probability by Car Ownership",
       x = "Car Ownership (HC10E)", y = "Pr(UB8 = 1)") +
  theme_bw(base_size = 12)

print(p_ed5a)
print(p_windex)
```

```{r}
#  4. Random effects (regional variation) 
re_df <- ranef(model_sign, condVar = TRUE)$region %>%
  tibble::rownames_to_column("region") %>%
  rename(intercept = `(Intercept)`) %>%
  mutate(
    se = sqrt(attr(ranef(model_sign, condVar = TRUE)$region, "postVar")[1, 1, ]),
    lower = intercept - 1.96 * se,
    upper = intercept + 1.96 * se
  )

p_re <- ggplot(re_df, aes(x = intercept, y = reorder(region, intercept))) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0.3, color = "steelblue") +
  geom_point(size = 3, color = "steelblue") +
  labs(
    title = "Regional Random Intercepts (±1.96 SE)",
    x     = "Random Intercept (log-odds scale)",
    y     = "Region"
  ) +
  theme_bw(base_size = 12)

print(p_re)


#  5. Random slopes by region (model_extra: windex5 | region_mics) 
re_slopes_df <- ranef(model_extra, condVar = TRUE)$region_mics %>%
  tibble::rownames_to_column("region") %>%
  rename(intercept = `(Intercept)`, slope_windex = windex5)

p_slopes <- ggplot(re_slopes_df, aes(x = intercept, y = slope_windex, label = region)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  geom_point(size = 3, color = "steelblue") +
  ggrepel::geom_text_repel(size = 3) +
  labs(
    title = "Random Intercepts vs. Random Slopes for Wealth Index by Region",
    x     = "Random Intercept",
    y     = "Random Slope (windex5)"
  ) +
  theme_bw(base_size = 12)

print(p_slopes)


#  ICC and R² across models — summary bar chart 
model_list <- list(
  "Null"        = model_null,
  "For Pooling" = model_for_pooling,
  "Full"        = model_full,
  "Significant" = model_sign,
  "Extra"       = model_extra
)

r2_df <- purrr::map_dfr(names(model_list), function(nm) {
  r2_vals <- r2(model_list[[nm]])
  icc_result <- tryCatch(
    performance::icc(model_list[[nm]])$ICC_adjusted,
    error = function(e) NA_real_
  )
  
  tibble::tibble(
    model        = nm,
    R2_marginal  = r2_vals$R2_marginal,
    R2_conditional = r2_vals$R2_conditional,
    ICC          = icc_result
  )
}) %>%
  tidyr::pivot_longer(-model, names_to = "metric", values_to = "value") %>%
  mutate(model = factor(model, levels = names(model_list)))

p_r2 <- ggplot(r2_df, aes(x = model, y = value, fill = metric)) +
  geom_col(position = "dodge") +
  scale_fill_manual(
    values = c("R2_marginal" = "#4575b4", "R2_conditional" = "#74add1", "ICC" = "#f46d43"),
    labels = c("ICC", "R² Conditional", "R² Marginal")
  ) +
  labs(
    title = "Model Fit Comparison: R² and ICC",
    x     = NULL,
    y     = "Value",
    fill  = NULL
  ) +
  ylim(0, 1) +
  theme_bw(base_size = 12) +
  theme(legend.position = "bottom")

print(p_r2)

```

# Model predictive performance comparison

```{r}
# Splitting data

uzb_df <- uzb_df %>%
  mutate(strata_var = paste0(region_mics, "_", UB8))

train_idx <- createDataPartition(uzb_df$strata_var, p = 0.75, list = FALSE)
train_df  <- uzb_df[train_idx, ]
test_df   <- uzb_df[-train_idx, ]


# Refit all models on training data
# Swapping uzb_df to train_df in each model call
# Example:
model_null_train <- update(model_null, data = train_df)
model_pool_train <- update(model_for_pooling, data = train_df)
model_full_train <- update(model_full, data = train_df)
model_sign_train <- update(model_sign, data = train_df)
model_extra_train <- update(model_extra, data = train_df)

# Predict test set
# allow.new.levels = TRUE handles regions in test not seen in train
predict_prob <- function(model, newdata) {
  predict(model, newdata = newdata, type = "response", allow.new.levels = TRUE)
}

probs <- list(
  Null = predict_prob(model_null_train, test_df),
  For_Pooling = predict_prob(model_pool_train, test_df),
  Full = predict_prob(model_full_train, test_df),
  Significant = predict_prob(model_sign_train, test_df),
  Extra = predict_prob(model_extra_train, test_df)
)


# Metrics

metrics_df <- purrr::map2_dfr(
  probs, names(probs),
  ~ compute_metrics(.x, as.numeric(as.character(test_df$UB8)), .y)
)

```

# Performance comparison table

```{r}
metrics_df %>%
  mutate(across(where(is.numeric), ~ round(.x, 3))) %>%
  gt::gt() %>%
  gt::tab_header(title = "Out-of-Sample Prediction Accuracy Comparison") %>%
  gt::fmt_number(columns = where(is.numeric), decimals = 3) %>%
  gt::cols_label(
    Model       = "Model",
    AUC         = "AUC",
    Accuracy    = "Accuracy",
    Sensitivity = "Sensitivity",
    Specificity = "Specificity",
    PPV         = "Precision (PPV)",
    F1          = "F1 Score",
    Brier       = "Brier Score",
    LogLoss     = "Log Loss"
  ) %>%
  gt::tab_style(
    style = gt::cell_fill(color = "#d4edda"),
    locations = gt::cells_body(
      rows = Accuracy == max(Accuracy)
    )
  )

# ROC curves plot (all models)
roc_list <- purrr::map2(probs, names(probs), function(p, nm) {
  roc_obj <- roc(as.numeric(as.character(test_df$UB8)), p, quiet = TRUE)
  data.frame(
    model       = paste0(nm, " (AUC=", round(auc(roc_obj), 3), ")"),
    sensitivity = roc_obj$sensitivities,
    specificity = roc_obj$specificities
  )
})

roc_df <- dplyr::bind_rows(roc_list)

p_roc <- ggplot(roc_df, aes(x = 1 - specificity, y = sensitivity, color = model)) +
  geom_line(linewidth = 0.9) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50") +
  scale_color_brewer(palette = "Set1") +
  labs(
    title = "ROC Curves — Out-of-Sample Performance",
    x     = "1 - Specificity (FPR)",
    y     = "Sensitivity (TPR)",
    color = NULL
  ) +
  theme_bw(base_size = 12) +
  theme(legend.position = "bottom")

print(p_roc)


# Calibration plot — predicted vs. observed probabilities
cal_df <- purrr::map2_dfr(probs, names(probs), function(p, nm) {
  data.frame(
    model    = nm,
    pred     = p,
    observed = as.numeric(as.character(test_df$UB8))
  )
})

# Bin predictions into deciles per model
cal_binned <- cal_df %>%
  dplyr::group_by(model) %>%
  dplyr::mutate(bin = ntile(pred, 10)) %>%
  dplyr::group_by(model, bin) %>%
  dplyr::summarise(
    mean_pred = mean(pred),
    mean_obs  = mean(observed),
    .groups   = "drop"
  )

p_cal <- ggplot(cal_binned, aes(x = mean_pred, y = mean_obs, color = model)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50") +
  geom_line(linewidth = 0.7, alpha = 0.8) +
  geom_point(size = 2) +
  scale_color_brewer(palette = "Set1") +
  labs(
    title = "Calibration Plot — Predicted vs. Observed Probabilities",
    x     = "Mean Predicted Probability",
    y     = "Mean Observed Proportion",
    color = NULL
  ) +
  theme_bw(base_size = 12) +
  theme(legend.position = "bottom")

print(p_cal)
```